Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2021-10-01 vertical-thermal-review.Rmd
"finished" canopy disturbance section

Variation in resistance and resilience among forests may drive different trajectories in the face of global climate change. 
For example, canopy trees of African rainforests appear to be more resilient to warming and drought than other tropical forests [@bennettResistanceAfricanTropical2021]. 
Species with smaller leaves and larger $g_s$ will be less vulnerable to temperature increase than species with larger leaves [@fauset_differences_2018; @songTreeSurfaceTemperature2020] and isoprene emitting species will have a selective advantage in a warming world [@taylorCapacityEmitIsoprene2019].

Canopy disturbance along with a trend towards younger forests and again reduced tree growth associated with warming [@mcdowellPervasiveShiftsForest2020] poses an increasing threat to the biodiversity of microrefugia that are otherwise buffered from climatic extremes [@scheffersMicrohabitatsReduceAnimal2013; @greiserHidingClimateCharacterizing2019]. 
Canopy structure affects understory species composition, which has been shown to shift under warming and canopy disturbance [@majasalmiImpactTreeCanopy2020; @bertrandCommentForestMicroclimate2020; @maesPlantFunctionalTrait2020]
In the understory, warming is disproportionately affecting the less thermally-adapted plant species, resulting in associated plant community thermophilization [@zellwegerForestMicroclimateDynamics2020; @duqueThermophilizationAdultJuvenile2015; @greiserHidingClimateCharacterizing2019].

*

These shifts in forest structure at large may be detectable by lidar and thermal remote sensing. If we know how metabolism maps to the thermal profile, then the observed forest structural changes can inform prediction of forest function.

@@

>>>>

Canopy disturbance poses an increasing threat to the biodiversity of microrefugia that are otherwise buffered from climatic extremes [@scheffersMicrohabitatsReduceAnimal2013; @greiserHidingClimateCharacterizing2019]. 
Canopy structure affects understory species composition, which has been shown to shift under warming and canopy disturbance [@majasalmiImpactTreeCanopy2020; @bertrandCommentForestMicroclimate2020; @maesPlantFunctionalTrait2020]
In the understory, warming is disproportionately affecting the less thermally-adapted plant species, resulting in associated plant community thermophilization [@zellwegerForestMicroclimateDynamics2020; @duqueThermophilizationAdultJuvenile2015; @greiserHidingClimateCharacterizing2019].
If such compositional shifts towards more thermally-adapted species fail to keep up with the pace of warming, the ecosystem-level resilience to canopy disturbance that is often provided by smaller trees [e.g., **REFS**, @millerReducedImpactLogging2011] will be destroyed, resulting in breakdown of canopy buffering and the potential state changes described above.

@@

As we have reviewed here,
